Predicting the 2024 U.S. Presidential Election: A Data-Driven Analysis of Swing State Outcomes

### 2024-11-04 paper/paper.qmd
minor edits

```{python}
#| echo: false
import pandas as pd
import matplotlib.pyplot as plt
import jupyter
from tabulate import tabulate
import numpy as np
```

# 1 Introduction
This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result. By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump victory in each state. This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.

@@

```{python}
#| ccho: false
#| label: Fig-1
#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538. The chart distinguishes between Democrat, Republican, and swing states, with a color-coded fill under each party’s cumulative curve. Key threshold of 270 electoral vote mark needed to secure a victory is diplayed. This visualization shows the democrats lead with 227 votes and the Republicans with 219 vot secured"
# Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot

@@

plt.figure(figsize=(20, 14))
plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
x_reversed = np.flip(x)
plt.step(x_reversed, cumulative_votes_reversed, where='post', color='red', linewidth=2, marker='o')

@@

# Adding threshold line and markers
plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
plt.axhline(270, color='purple', linestyle='--', linewidth=2)
plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
plt.xlabel('Electoral College by State', fontsize=12)
plt.ylabel('Total Votes', fontsize=12)
plt.title('Cumulative Electoral Votes by State', fontsize=25)
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()

@@

The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on its own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@brick2013Nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.

@@

## what is included in the poll?

## Methodology and why?

# References

>>>>

```{python}
#| echo: false
import pandas as pd
import matplotlib.pyplot as plt
import jupyter
from tabulate import tabulate
import numpy as np
```
\newpage
# 1 Introduction
This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result. By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump victory in each state. This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.

@@

```{python}
#| ccho: false
#| label: Fig-1
#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538. The chart distinguishes between Democrat, Republican, and swing states, with a color-coded fill under each party’s cumulative curve. Key threshold of 270 electoral vote mark needed to secure a victory is diplayed. This visualization shows the democrats lead with 227 votes and the Republicans with 219 votes secured."
# Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot

@@

plt.figure(figsize=(18, 14))
stop = 320
plt.step(x[:stop], cumulative_votes[:stop], where='post', color='navy', linewidth=2, marker='o')
x_reversed = np.flip(x)
plt.step(x_reversed, cumulative_votes_reversed, where='post', color='red', linewidth=2, marker='o')

@@

# Adding threshold line and markers
plt.xticks(range(len(states)), states, rotation=90, fontsize=13)
plt.axhline(270, color='purple', linestyle='--', linewidth=2)
plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
plt.xlabel('Electoral College by State', fontsize=15)
plt.ylabel('Total Votes', fontsize=15)
plt.title('Cumulative Electoral Votes by State', fontsize=25)
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()

@@

The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on its own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@brick2013nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.

@@

\newpage
# References
